Application of Micro-Rotors to Induce Crystal Nucleation: Challenges, Takeaways, and Future Prospects\\ PhD Thesis

### 2025-07-04 Latex_files/Thesis/chapters/chapter 6/main.tex
fleshing out QPD chapter and edits to new chp 3&6

Where $S$ is the supersaturation from eq.~\eqref{eq:supersaturation} 
$v$ is the approximate volume of an individual molecule, 
$k_B$ is the Boltzmann constant, and $\sigma_{inf}$ is the 
interfacial tension of the bulk solution. This assumes that 
the nucleus will have a spherical morphology so that the 
surface tension $\sigma_{inf}$ is a scalar value. Looking at 
\eqref{eq:CNT} suggests that there must be some critical size 
$r$ where the free energy gain from the nucleus exceeds the 
surface tension of the surrounding fluid. This is reflected in 
fig.~\ref{fig:free_energy} where we plot the free energy of the 
system against nucleus size. This reveals a critical size above 
which the gain in free energy exceeds the interfacial tension. 
Furthermore fig.~\ref{fig:free_energy} shows how increasing the 
supersaturation of the system reduces said barrier. 
\begin{figure}[h!]
	\centering
	\includegraphics[width=\linewidth]{Free_Energy_Diagram.png}
	\caption{Free energy diagram of a newly formed nucleus according 
		to the Classical Nucleation Theory. The total free energy (blue)
		is due to the competition between the volume free energy gain
		(green) and the interfacial free energy cost (orange). Dotted
		lines are for a higher supersaturation than the solid lines,
		the interfacial energy cost is independent of supersaturation.
		Diagram was constructed in python using a naive assumption 
		of $v=1\times 10^-12\ m^3$ and $\sigma_{inf}=0.0001\ J/m^2$}
	\label{fig:free_energy}
\end{figure}

@@

One of the key conclusions is the need for the development of 
better \textit{in situ} techniques for characterizing and 
imaging the nucleation pathways. Understanding which specific 
pathway is unfolding under a certain set of experimental 
parameters would allow one to develop a map of the thermodynamic 
landscape available to a pre-nucleation system. The challenge 
therein lies in developing experimental methods by which one can 
reproduce similar nucleation events and in turn use 
characterisation techniques to identify the pathway taken. The 
method not only needs to be repeatable for a single set of system 
parameters, but also needs to be flexible enough to replicate 
results regardless of the system parameters (i.e. varying 
supersaturation, temperature, and solute choice). There have 
been several different methods deployed to try and observe 
crystallisation at a microscopic scale.

\subsection{Computer Simulations}
While not a direct observation of nucleation events, computer
simulations have proved an invaluable tool for studying and 
testing predictions about the pre-nucleated solution. By 
comparing the simulative nucleation rate to the expected 
rate given by current models, researchers can test newly 
developed theories and glean information about the microscopic
parameters of a system.

The main challenge facing computer simulations is the issue
of limited scope. Molecular dynamic simulations are constrained
by their choice of time and length scale, for example simulations
in a canonical ensemble face issues where the chemical potential
is depleted to the point that the driving force for crystallisation
is halted \cite{DuranOlivencia2015, Finney2023}, this can be 
addressed but requires an extensive increase in computational 
resources to maintain the chemical potential equilibria. Like wise
in most cases the choice of time scale is a crucial factor, shorter
time scales provide a more accurate evolution of the solution; 
however, in some cases the time scale magnitude for crystallisation
is often far greater than the time scale used in the simulation. This
has led to more advanced sampling techniques that allow for longer 
time scales which has provided insights into the pre-nucleation 
solution stage and formation of pre-nucleated clusters \cite{Finney2023}.

The study of the pre-nucleated solution and the evolution of 
precursors has been aided greatly by the use of simulative 
studies. Early on it was shown in colloidal crystallisation 
that the structure of the precursor was time dependent; 
colloidal simulations showed that regardless of the molecular
interactions between colloids the atomic packing pattern was 
predominately hexagonal centred, over time the packing
became dominated by body centred packing in the case 
where the interactions were longer reaching \cite{Tan2013}. 
They also confirmed that precursor formation was driven not
by local density variations, but instead was driven by the 
local bond order of the colloidal precursors \cite{Tan2013}, 
providing strong evidence that the colloidal pre-nucleation 
clusters undergo rearrangement prior to nucleation. Further
developments in $NaCl$ simulations have shown that by varying
the density fluctuations the precursor structure could either
be amorphous, the classical rock-salt structure, or a new 
structure similar to wurtzite. While this result is difficult
to confirm it provides a important conclusion, that simulations
can hint at new nucleation pathways that have not been confirmed
experimentally. This means that computer simulations can not
be merely used to confirm theories, but can be used to examine
and search for novel results.

Overall computer simulations make up a core area of research 
in the study of nucleation. The technique is however limited,
often struggling to examine nucleation over longer time scales
or larger number of particles. While sampling methods and 
enhanced simulation techniques has alleviated this somewhat 
they are still far from being able to describe the behaviour 
of larger molecules undergoing nucleation.

\subsection{Transmission Electron Microscope (TEM)}
One of the more well known techniques used for direct observations 
of crystallisation dynamics is using transmission electron 
microscopy (TEM). The basic working principle involves creating a 
focused beam of electrons that are directed onto a target sample.
Due to the wave-particle duality, the electrons can be treated 
as a unified beam of light, whose wavelength is dependent on the 
kinetic energy of the electrons. The wavelength of an individual 
electron is given by \cite{Williams2009}:
\begin{align}
	\lambda_e = \frac{h}{\sqrt{2mE(1+\frac{E}{2mc^2})}}
\end{align}

\noindent
When incident on a sample the wave is scattered by the sample 
and can then be detected by an imaging screen that measures 
the difference in intensity between the scattered and 
transmitted beams. The advantage of TEM is that one can bypass
the diffraction limit imposed by visible light, the maximum 
resolution that can be achieved by a microscope is given by
\cite{Champness2020}:
\begin{align}
	d = \frac{\lambda}{2NA}
\end{align}

\noindent
where NA is the numerical aperture of the objective lens, and
$\lambda$ is the wavelength of light used to illuminate a sample.
For a visible light LED with an average wavelength of $550~nm$, 
the upper limit for an objective lens (NA of 1.2) is around 
$230~nm$ whereas a similar lens focusing an electron beam 
with an average kinetic energy of $100~keV$ will have a maximum 
resolution of just over $4~pm$ \cite{Champness2020}. There is
an upper limit to the maximum resolution possible as electron 
lens are far inferior to what modern optics can achieve 
\cite{Champness2020} so there is a trade off between quality and 
resolution.

TEM technology has proven invaluable to the study of nucleation
at an atomic level. The samples are often only a few hundred 
atoms thick in order to optimise the imaging results. Much of 
the supporting evidence for non-classical nucleation has been 
demonstrated via TEM \cite{Ye2023}. For example when Cao 
\textit{et al} observed the formation of amorphous atomic 
cluster using $\gamma-$Fe,Au, and Re; all three of which 
underwent continuous transformation before eventually 
reaching an final ordered state \cite{Cao2020}. The amorphous 
cluster did not immediately transition to an ordered structure, 
instead fluctuating somewhat randomly until eventually settling 
into a crystalline structure \cite{Cao2020}. This was confirmed 
both in solid growth and liquid growth conditions \cite{Cao2020, 
	Loh2016}, suggesting that the non-classical nucleation path 
dominates in all cases of metal crystal formation \cite{Ye2023}. 
Confirmation is necessary for different solute choices, as in 
the case of organic solids it is believed that non-classical 
nucleation only dominates in low supersaturations whereas 
classical nucleation is dominate at higher supersaturations.

TEM analysis has not been extended to the study of organic 
crystals, proteins, or simple salts at an atomic level 
however. The main reason being that organic materials can be 
damaged by the electron beam and therefore need to be frozen 
prior to being imaged \cite{Champness2020}. This was used to 
study the polymorphic time dependence of glycine crystals, by 
varying the time prior to freezing Broadhurst \textit{et al} 
showed that $\beta$ glycine dominates initially but $\alpha$-
glycine is more stable after a longer period, $\gamma$-glycine 
was only found when allowed to dry over an hour on a glass slide \cite{Broadhurst2020}. While this was invaluable for showing 
the time scales with which polymorphs form, it doesn't provide 
much information on the kinetics of crystal growth as these 
were snap shots and not direct observations of the nucleation 
process.

Overall TEM analysis is invaluable for the study of atomic 
crystal growth, the secondary scattering from electron 
beams also allows for chemical and structural information 
to be gleaned from the target. This has given strong evidence
for the non-classical nucleation theory, particularly in 
metal crystal formation. Due to the high energy of the 
electrons, organic materials have not received the same
focus outside of cryoTEM analysis. Non-invasive methods
are therefore necessary to gain an understanding of the
crystal growth and formation of organic crystals, 
particularly when both classical and multi-step nucleation 
can occur.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Optical Tweezer Equipment}
In general, all optical tweezers require a laser driver, a focusing 
microscope objectives, a position controller, and position detector 
\cite{Gieseler2020}. The laser used for this project was a $1064\ 
nm$ near infrared laser - provided by CNI Lasers – that was focused 
by a Nikon 100x oil immersion lens. The choice of an oil immersion 
lens is important as the optical oil used prevents a loss of focus 
when used on a glass cover slip. Now, experimental work has shown 
that the trapping efficiency increases with beam diameter up until 
it exceeds $\frac{2}{3}D_{obj}$ \cite{kim2003dependence} where 
$D_{obj}$ is the diameter of the objective aperture. To expand the 
beam front we utilise a Galilean beam expansion arrangement 
(indicated by $f_1$, and $f_2$ in Fig.~\ref{fig:setup}) as recommended 
for high power laser applications. In our initial experiments the 
beam expansion provides a $4\times$ magnification. Whereas in later 
experiments we utilised a galvano-mirror the beam expansion is 
$3\times$ and then the 4f correlator (see \ref{sec:4f_correlation})
provides a further $1.25\times$ magnification (using $f_3$ and $f_4$) 
- the magnification is given by.
\begin{align}
	\frac{D_2}{D_1} = \frac{f_2}{f_1}
\end{align}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Fourier Optics and 4f correlators}
\label{sec:4f_correlation}
As shown in \ref{fig:setup}, after the galvano-mirror there are a pair
of focal lenses ($f_3$ and $f_4$) that do not seem to serve a clear
purpose. These are in fact crucial for the operation of the Galvano, the
pair of them can be more accurately called a 4f correlator.

@@

Due to Van der Waal's forces some of the microspheres were stuck 
together, fortunately individual sphere's were still present. 
Individual microspheres were trapped and their rotation rate was 
determined by looking at the peak frequency component of the 
collected power spectrum. Due to their anisotropic scattering the 
QPD signal will oscillate periodically. After taking the Fourier
transform of the QPD signal the power spectrum has distinctive
peaks at integer multiples of the rotational frequency. The largest 
peak is often twice the rotational frequency which is due to the 
polarization modulation of the scattered field, the first peak 
is due purely to the rotation of vaterite sphere \cite{Monteiro2018}. 
An example of this is behaviour is shown in fig.~\ref{fig:tdat_fdat}.

@@

The optimal shear rate predicted by \cite{Debuysschere2023} is 
plotted on Fig.~\ref{fig:vaterite_shear}(b), the outer dotted 
lines represent the point when the nucleation rate is less than 
90\% of the maximum. We can visualise the fluid around the 
rotating sphere by dividing it up into radial sections. Assuming 
the sphere is rotating sufficiently quickly there are three 
sections of concern, as demonstrated by fig.~\ref{fig:shear_diagram}. 
The fluid close to the sphere experiences a shear rate that 
is far greater than the optimal shear rate, in which case 
nucleation is suppressed \cite{Mura2016}. Secondly, there is 
a region of fluid that experiences a shear rate close to the 
optimal value ($\pm 5\%$) where the nucleation rate is enhanced 
significantly compared to the bulk fluid. Lastly, beyond the 
optimal region the shear rate drops off exponentially in which 
case the nucleation rate is barely different compared to the 
bulk fluid.
\begin{figure}[h!]
	\centering
	\includegraphics[width=0.875\linewidth]{shear_rate_diagram.pdf}
	\caption{Diagram demonstrating how fluid shear rate varies 
	radially from a rotating sphere (rotation rate = $f_{rot}$). 
	(a) the shear rate exceeds the optimal shear rate 
	($\dot{\gamma}_{opt}$) and suppresses the nucleation rate. 
	(b) shear rate is close to the optimal amount, enhancing the 
	nucleation rate. (c) shear rate is far below the optimal 
	amount and the nucleation rate is comparable to the bulk fluid.}
	\label{fig:shear_diagram}
\end{figure}

@@

Estimating the distance from the trap focus to the cover slip
is difficult to measure directly, often requiring one fits a 
complex Lorentzian curve to the recorded power spectra in order 
to measure the fluid drag from being in close proximity to the 
cover slip \cite{BergSoerensen2004}. We estimate that the 
vaterite must be at least $15-20\mu\ m$ from the cover slip for 
us to reliably trap them. The contact angle between the cover 
slip and solution was previously measured in \cite{Flannigan2023},
while being loosely dependent on the supersaturation the 
measured contact angle for supersaturated solutions was between 
$30^{\circ}$ and $35^{\circ}$. This is shown in fig.\ref{fig:3.9}.
\begin{figure}[h!]
	\centering
	\includegraphics[width=\linewidth]{vaterite_diagram.pdf}
	\caption{Diagram of optical trapping set up for rotating 
		birefringent particles in a supersaturated solution. 
		Left: side view of the trapping set up showing the 
		location of the trap focus at the edge of the droplet 
		of a supersaturated solution. The approximate contact 
		angle (calculated by \cite{Flannigan2023}) is used to 
		provide a scale of how close a trapped particle could 
		be located. Right: top down view of the glycine droplet 
		with a trapped birefringent particle shown close to 
		edge of the trap. As the particle rotates the drag 
		force from the surrounding fluid generates a flow field 
		around itself (see Eq.~\ref{eq:birefringent_speed}).}
	\label{fig:3.9}
\end{figure}

@@

It is suspected that trapping is much harder at the 
interface due to increased surface tension and unpredictable 
scattering forces. The closest we could trap a microsphere to 
the droplet edge was in the range of $5-10\mu\ m$, at that 
distance the fluid flow is so low that even the presence of a 
hard boundary would be insufficient for shearing the fluid. 
Furthermore, as is evident in Table~\ref{tab:Nucleation}, 
the rotation rate drops off significantly with increased 
supersaturation, due to higher fluid viscosities. While in 
theory a sufficiently focused laser could rotate any 
microsphere to a fast enough to reach the shear rate 
predicted by \cite{Debuysschere2023} the localised intensity 
would be so large that even using $D_2O$ would see a significant 
increase in temperature.

It is not impossible that fluid shearing could be used in the 
future to localise nucleation; but from these results, using 
individual micro-rotors is not an appropriate method due to
two key factors. Firstly, the area of influence is far too 
small to see any noticeable increase in the nucleation rate 
this is demonstrated most clearly in fig.~\ref{fig:vaterite_shear}. 
Due to the exponential decay in the expected shear rate it is 
very difficult trying to shear a large volume of fluid. Even 
if the optimal shear rate at a micro-level was an order of 
magnitude less than what was predicted by \cite{Debuysschere2023}, 
only a small volume of fluid would even experience that shear 
rate. In order to properly understand the microscopic effects 
of fluid shear a method of localised shearing over a larger 
fluid volume is necessary.

@@

If multiple micro-rotors could be trapped in close proximity 
to one another they could create a large region of fluid where 
nucleation is more likely than the bulk fluid. Because optical
torque is not contingent on the fluid properties the only 
limiting factor would be the total angular momentum transferred 
to each particle. Micro-rotors have been created that allow for 
precise control of suspended micro-particles \cite{Butaite2019} 
and could potentially be used to generate sufficient shearing. 
However these could not be used in this project as we lacked the 
necessary hardware to form multiple gradient traps.  
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Nucleation with a Stationary and Moving Beam}
As mentioned previously, shearing via optical rotation did 
not result in any localised nucleation events even while 
in the proximity of the droplet edge. An alternative approach
was suggested, using a galvano mirror to move a particle 
quickly through a supersaturated solution. Preliminary 
calculations suggested a particle moving rapidly through 
a fluid could produce significantly greater shear rates 
(see \emph{Appendix X.X} for breakdown) than a rotating 
particle. While a particle is within the optical trap the 
gradient forces are significantly reduced meaning simply 
trapping a particle is insufficient for inducing nucleation \cite{Flannigan2023}. This is why we did not see any 
nucleation events while rotating the vaterite, there is 
not a large gradient force to draw in material.

@@

\subsubsection{Choice of Crystal growth rate units}
The growth rate of a crystal can be described using different
units depending on the situation. For example, seed crystals 
often use growth units in the form of \textit{length per unit
time}. This can either be used to describe the overall length
of the longest edge of a crystal, or to describe how individual
faces grow separately. This works well for crystals with a 
clear morphology, in regards to laser induced nucleation the 
crystal growth is heavily influenced by the local fluid 
conditions meaning tracking the length of an individual face 
is challenging. Instead we utilise image analysis software 
such as imageJ to measure the area of the crystal after each 
frame to get a rough estimate of the crystal growth rate in 
units of \textit{length$^2$ per unit time}. With the addition 
of the contact angle information from \cite{Flannigan2023} we 
can approximate the crystal height based on how close it is to 
the droplet edge.

@@

Shown below in Fig.~\ref{fig:eliptical_beam_1} where 
we have the laser focus moving in a small elliptical 
pattern. While nothing is seen directly entering the 
focus a nucleus forms close to the droplet edge, 
unlike in fig.~\ref{fig:stationary_beam} the crystal 
does not grow out from the focal point evenly. Due to 
the galvano mirror, the crystal is simultaneously 
being moved by and growing around the focal point of 
the trap. Because of this the crystal nucleus lacks 
a clear morphology at first. Until roughly $20\ s$ 
the crystal reaches a almost prismatic structure, 
with further irradiation increasing the size.
\begin{figure}[h!]
	\centering
	\includegraphics[width=0.8\linewidth]{frames_eliptical_beam.pdf}
	\caption{Frames from a longer video depicting the growth 
		of a nucleus using a moving beam. Beam power is kept 
		at 700 mW and the supersaturation was estimated at 
		S=0.86. Initially the crystal shape is amorphous (a) 
		but eventually reaches a more regular shape (b). This crystal is still influenced by the optical trap as 
		even when not directly irradiated by the laser the 
		crystal rotates between (c) and (d). When the laser 
		is focused on a corner the crystal growth is localised 
		to that region, resulting in an elongated section 
		forming between frames (e) and (f).}
	\label{fig:eliptical_beam_1}
\end{figure}

Interestingly the galvano-mirror allows the trap to 
impart a slight torque on the crystal, as shown in fig.~\ref{fig:eliptical_beam_1}(c) and (d), where even 
though the crystal is not directly in the trap focus 
it rotates in the $x-y$ plane and gets trapped again 
at a corner. The rotation could not be due to fluid 
flow close to the surface of the crystal as the 
dipole moment of individual water molecules is too 
small to be influenced by an optical trap. In figs.\ref{fig:eliptical_beam_1}(e) and (f), the 
crystal growth becomes localised to the corner. The 
area growth rate between figures \ref{fig:eliptical_beam_1}
(a) and (d) was approximated using imageJ at $45.03\ 
\mu m^2 /min$, where as between figures 
\ref{fig:eliptical_beam_1}(e) and (f) the growth rate 
at that particular edge was estimated at $42.10\ \mu m^2/min$.

>>>>

Where $S$ is the supersaturation from eq.~\eqref{eq:supersaturation} 
$v$ is the approximate volume of an individual molecule, 
$k_B$ is the Boltzmann constant, and $\sigma_{inf}$ is the 
interfacial tension of the bulk solution. This assumes that 
the nucleus will have a spherical morphology so that the 
surface tension $\sigma_{inf}$ is a scalar value. Looking at 
\eqref{eq:CNT} suggests that there must be some critical size 
$r$ where the free energy gain from the nucleus exceeds the 
surface tension of the surrounding fluid. This is reflected in 
Fig.~\ref{fig:free_energy} where we plot the free energy of the 
system against nucleus size. This reveals a critical size above 
which the gain in free energy exceeds the interfacial tension. 
Furthermore Fig.~\ref{fig:free_energy} shows how increasing the 
supersaturation of the system reduces said barrier. 
\begin{figure}[h!]
	\centering
	\includegraphics[width=\linewidth]{Free_Energy_Diagram.png}
	\caption{Free energy diagram of a newly formed nucleus according 
		to the Classical Nucleation Theory. The total free energy (blue)
		is due to the competition between the volume free energy gain
		(green) and the interfacial free energy cost (orange). Dotted
		lines are for a higher supersaturation than the solid lines,
		the interfacial energy cost is independent of supersaturation.
		Diagram was constructed in python using a naive assumption 
		of $v=1\times 10^-12\ m^3$ and $\sigma_{inf}=0.0001\ J/m^2$}
	\label{fig:free_energy}
\end{figure}

@@

One of the key conclusions is the need for the development of 
better \textit{in situ} techniques for characterizing and 
imaging the nucleation pathways. Understanding which specific 
pathway is unfolding under a certain set of experimental 
parameters would allow one to develop a map of the thermodynamic 
landscape available to a pre-nucleation system. The challenge 
therein lies in developing experimental methods by which one can 
reproduce similar nucleation events and in turn use 
characterisation techniques to identify the pathway taken. The 
method not only needs to be repeatable for a single set of system 
parameters, but also needs to be flexible enough to replicate 
results regardless of the system parameters (i.e. varying 
supersaturation, temperature, and solute choice).

This was one of the initial goals of the PhD, to localise nucleation
using optical tweezing. While tweezing by itself can induce nucleation
the exact mechanism that causes it is still unclear. Optical interactions
with sub-wavelength particles are not possible so only a two-step 
nucleation event can be occur via optical tweezing. In order to accurately 
characterise the nucleation pathway taken at a given supersaturation, 
a method is needed that is not dependent on the pathway taken. This is 
what lead us to consider localised shearing, while still favoured by 
two-step nucleation it is still possible for the classical nucleation 
pathway to occur, given that the free energy barrier is low enough to 
be traversed via localised shearing.  
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Optical Tweezer Equipment}
In general, all optical tweezers require a laser driver, a focusing 
microscope objectives, a position controller, and position detector 
\cite{Gieseler2020}. The laser used for this project was a $1064\ 
nm$ near infrared laser - provided by CNI Lasers – that was focused 
by a Nikon 100x oil immersion lens. The choice of an oil immersion 
lens is important as the optical oil used prevents a loss of focus 
when used on a glass cover slip. Now, experimental work has shown 
that the trapping efficiency increases with beam diameter up until 
it exceeds $\frac{2}{3}D_{obj}$ \cite{kim2003dependence} where 
$D_{obj}$ is the diameter of the objective aperture. To expand the 
beam front we utilise a Galilean beam expansion arrangement 
(indicated by $f_1$, and $f_2$ in Fig.~\ref{fig:setup}) as recommended 
for high power laser applications. In our initial experiments the 
beam expansion provides a $4\times$ magnification. Whereas in later 
experiments we utilised a galvano-mirror the beam expansion is 
$3\times$ and then the 4f correlator (see \ref{sec:4f_correlation})
provides a further $1.25\times$ magnification (using $f_3$ and $f_4$) 
- the magnification is given by.
\begin{align}
	\frac{D_2}{D_1} = \frac{f_2}{f_1}
\end{align}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Fourier Optics and 4f correlators}
\label{sec:4f_correlation}
As shown in Fig.~\ref{fig:setup}, after the galvano-mirror there are a pair
of focal lenses ($f_3$ and $f_4$) that do not seem to serve a clear
purpose. These are in fact crucial for the operation of the Galvano, the
pair of them can be more accurately called a 4f correlator.

@@

Due to Van der Waal's forces some of the microspheres were stuck 
together, fortunately individual sphere's were still present. 
Individual microspheres were trapped and their rotation rate was 
determined by looking at the peak frequency component of the 
collected power spectrum. Due to their anisotropic scattering the 
QPD signal will oscillate periodically. After taking the Fourier
transform of the QPD signal the power spectrum has distinctive
peaks at integer multiples of the rotational frequency. The largest 
peak is often twice the rotational frequency which is due to the 
polarization modulation of the scattered field, the first peak 
is due purely to the rotation of vaterite sphere \cite{Monteiro2018}. 
An example of this is behaviour is shown in Fig.~\ref{fig:tdat_fdat}.

@@

The optimal shear rate predicted by \cite{Debuysschere2023} is 
plotted on Fig.~\ref{fig:vaterite_shear}(b), the outer dotted 
lines represent the point when the nucleation rate is less than 
90\% of the maximum. We can visualise the fluid around the 
rotating sphere by dividing it up into radial sections. Assuming 
the sphere is rotating sufficiently quickly there are three 
sections of concern, as demonstrated by Fig.~\ref{fig:shear_diagram}. 
The fluid close to the sphere experiences a shear rate that 
is far greater than the optimal shear rate, in which case 
nucleation is suppressed \cite{Mura2016}. Secondly, there is 
a region of fluid that experiences a shear rate close to the 
optimal value ($\pm 5\%$) where the nucleation rate is enhanced 
significantly compared to the bulk fluid. Lastly, beyond the 
optimal region the shear rate drops off exponentially in which 
case the nucleation rate is barely different compared to the 
bulk fluid.
\begin{figure}[h!]
	\centering
	\includegraphics[width=0.875\linewidth]{shear_rate_diagram.pdf}
	\caption{Diagram demonstrating how fluid shear rate varies 
	radially from a rotating sphere (rotation rate = $f_{rot}$). 
	(a) the shear rate exceeds the optimal shear rate 
	($\dot{\gamma}_{opt}$) and suppresses the nucleation rate. 
	(b) shear rate is close to the optimal amount, enhancing the 
	nucleation rate. (c) shear rate is far below the optimal 
	amount and the nucleation rate is comparable to the bulk fluid.}
	\label{fig:shear_diagram}
\end{figure}

@@

Estimating the distance from the trap focus to the cover slip
is difficult to measure directly, often requiring one fits a 
complex Lorentzian curve to the recorded power spectra in order 
to measure the fluid drag from being in close proximity to the 
cover slip \cite{BergSoerensen2004}. We estimate that the 
vaterite must be at least $15-20\ \mu m$ from the cover slip for 
us to reliably trap them. The contact angle between the cover 
slip and solution was previously measured in \cite{Flannigan2023},
while being loosely dependent on the supersaturation the 
measured contact angle for supersaturated solutions was between 
$30^{\circ}$ and $35^{\circ}$. This is shown in Fig.\ref{fig:3.9}.
\begin{figure}[h!]
	\centering
	\includegraphics[width=\linewidth]{vaterite_diagram.pdf}
	\caption{Diagram of optical trapping set up for rotating 
		birefringent particles in a supersaturated solution. 
		Left: side view of the trapping set up showing the 
		location of the trap focus at the edge of the droplet 
		of a supersaturated solution. The approximate contact 
		angle (calculated by \cite{Flannigan2023}) is used to 
		provide a scale of how close a trapped particle could 
		be located. Right: top down view of the glycine droplet 
		with a trapped birefringent particle shown close to 
		edge of the trap. As the particle rotates the drag 
		force from the surrounding fluid generates a flow field 
		around itself (see Eq.~\ref{eq:birefringent_speed}).}
	\label{fig:3.9}
\end{figure}

@@

It is suspected that trapping is much harder at the 
interface due to increased surface tension and unpredictable 
scattering forces. The drag coefficient is estimated to increase
by 10\% when approaching a flat boundary \cite{Yao2009} so approaching 
an angled surface would see an even greater increase.  The 
closest we could trap a microsphere to the droplet edge was 
in the range of $5-10\ \mu m$, at that distance the fluid flow 
is so low that even the presence of a hard boundary would be 
insufficient for shearing the fluid. Furthermore, as is evident 
in Table~\ref{tab:Nucleation}, the rotation rate drops off 
significantly with increased supersaturation, due to higher 
fluid viscosities. Solution viscosity has been shown to increase 
with a slight addition of glycine \cite{Patyar2020}, though as 
best as we can find there have been no studies of solution 
viscosity when the concentration reaches saturation levels. 
One would suspect that when $S\geq1$ the chemical potential 
driving the solute molecules to nucleate would also somewhat 
resist the velocity field leading to an additional increase in 
viscosity.

While in theory a sufficiently focused laser could rotate any 
microsphere to a fast enough to reach the shear rate predicted 
by \cite{Debuysschere2023} the localised intensity would be so 
large that even using $D_2O$ would see a significant increase 
in temperature.

It is not impossible that fluid shearing could be used in the 
future to localise nucleation; but from these results, using 
individual micro-rotors is not an appropriate method due to
two key factors. Firstly, the area of influence is far too 
small to see any noticeable increase in the nucleation rate 
this is demonstrated most clearly in Fig.~\ref{fig:vaterite_shear}. 
Due to the exponential decay in the expected shear rate it is 
very difficult trying to shear a large volume of fluid. Even 
if the optimal shear rate at a micro-level was an order of 
magnitude less than what was predicted by \cite{Debuysschere2023}, 
only a small volume of fluid would even experience that shear 
rate. In order to properly understand the microscopic effects 
of fluid shear a method of localised shearing over a larger 
fluid volume is necessary.

@@

If multiple micro-rotors could be trapped in close proximity 
to one another they could create a large region of fluid where 
nucleation is more likely than the bulk fluid. Because optical
torque is not contingent on the fluid properties the only 
limiting factor would be the total angular momentum transferred 
to each particle. Micro-rotors have been created that allow for 
precise control of suspended micro-particles \cite{Butaite2019} 
and could potentially be used to generate sufficient shearing. 
However these could not be used in this project as we lacked the 
necessary hardware to form multiple gradient traps.  
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Nucleation with a Stationary and Moving Beam}
As mentioned previously, shearing via optical rotation did 
not result in any localised nucleation events even while 
in the proximity of the droplet edge. An alternative approach
was suggested, using a galvano mirror to move a particle 
quickly through a supersaturated solution. Preliminary 
calculations suggested a particle moving rapidly through 
a fluid could produce significantly greater shear rates 
(see Appendix\ref{sec.shear_rate-calcs} for breakdown) 
than a rotating particle. While a particle is within the 
optical trap the gradient forces are significantly reduced 
meaning simply trapping a particle is insufficient for 
inducing nucleation \cite{Flannigan2023}. This is why we did 
not see any nucleation events while rotating the vaterite, 
there is not a large gradient force to draw in material.

@@

\subsection{Choice of Crystal growth rate units}
The growth rate of a crystal can be described using different
units depending on the situation. For example, seed crystals 
often use growth units in the form of \textit{length per unit
time}. This can either be used to describe the overall length
of the longest edge of a crystal, or to describe how individual
faces grow separately. This works well for crystals with a 
clear morphology, in regards to laser induced nucleation the 
crystal growth is heavily influenced by the local fluid 
conditions meaning tracking the length of an individual face 
is challenging. Instead we utilise image analysis software 
such as imageJ to measure the area of the crystal after each 
frame to get a rough estimate of the crystal growth rate in 
units of \textit{length$^2$ per unit time}. With the addition 
of the contact angle information from \cite{Flannigan2023} we 
can approximate the crystal height based on how close it is to 
the droplet edge.

@@

Shown below in Fig.~\ref{fig:eliptical_beam_1} where 
we have the laser focus moving in a small elliptical 
pattern. While nothing is seen directly entering the 
focus a nucleus forms close to the droplet edge, 
unlike in Fig.~\ref{fig:stationary_beam} the crystal 
does not grow out from the focal point evenly. Due to 
the galvano mirror, the crystal is simultaneously 
being moved by and growing around the focal point of 
the trap. Because of this the crystal nucleus lacks 
a clear morphology at first. Until roughly $20\ s$ 
the crystal reaches a almost prismatic structure, 
with further irradiation increasing the size.
\begin{figure}[h!]
	\centering
	\includegraphics[width=0.8\linewidth]{frames_eliptical_beam.pdf}
	\caption{Frames from a longer video depicting the growth 
		of a nucleus using a moving beam. Beam power is kept 
		at 700 mW and the supersaturation was estimated at 
		S=0.86. Initially the crystal shape is amorphous (a) 
		but eventually reaches a more regular shape (b). This crystal is still influenced by the optical trap as 
		even when not directly irradiated by the laser the 
		crystal rotates between (c) and (d). When the laser 
		is focused on a corner the crystal growth is localised 
		to that region, resulting in an elongated section 
		forming between frames (e) and (f).}
	\label{fig:eliptical_beam_1}
\end{figure}

Interestingly the galvano-mirror allows the trap to 
impart a slight torque on the crystal, as shown in Fig.~\ref{fig:eliptical_beam_1}(c) and (d), where even 
though the crystal is not directly in the trap focus 
it rotates in the $x-y$ plane and gets trapped again 
at a corner. The rotation could not be due to fluid 
flow close to the surface of the crystal as the 
dipole moment of individual water molecules is too 
small to be influenced by an optical trap. In Figs.\ref{fig:eliptical_beam_1}(e) and (f), the 
crystal growth becomes localised to the corner. The 
area growth rate between Figures \ref{fig:eliptical_beam_1}
(a) and (d) was approximated using imageJ at $45.03\ 
\mu m^2 /min$, where as between Figures~.\ref{fig:eliptical_beam_1}
(e) and (f) the growth rate at that particular edge was estimated 
at $42.10\ \mu m^2/min$.

Nucleation in undersaturated conditions has been 
reported previously in $D_2O$ \cite{Rungsimanon2010} 
and $H_2O$ \cite{Flannigan2023}, though not involving 
a moving beam. This modification allows for the crystal 
growth to be localised to a specific region of the bulk 
crystal whereas with a stationary beam there is no 
control over the crystal morphology. In fact this allows 
for a much finer control over the exact shape of the 
crystal nucleus, in some cases allowing for growth out
of the viewing plane as shown in Fig.~\ref{fig:out_of_plane}.
\begin{figure}[h!]
	\centering
	\includegraphics[width=\linewidth]{out_of_plane.pdf}
	\caption{Interesting crystal growth on the surface of an
	existing crystal. Initially the crystal surface seems 
	flat (a) but eventually a different crystal structure 
	grows out off of the surface (b).}
	\label{fig:out_of_plane}
\end{figure}
